```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here's some of the code from [GwR chapter 4 section 2: Spatial operations on vector data](https://geocompr.robinlovelace.net/spatial-operations.html#spatial-vec), for you to follow along with as we go through it in class.

## Prerequisites

```{r libraries}
library(tidyverse)
library(sf)
library(spData)
library(tmap)
```

## Spatial operations on vector data

### Spatial subsetting

```{r}
# subset nz by attribute
#
canterbury = nz %>% filter(Name == "Canterbury")

# subset nz_height by polygon intersection
#
canterbury_height = nz_height[canterbury, , op = st_intersects ]
```

```{r}
tmap_mode("view")
tm_basemap("OpenStreetMap") +
  tm_shape(canterbury) + tm_borders() +
  tm_shape(canterbury_height) + tm_dots(col = "red", size = 0.05)

# use geometry as a stencil to pull stuff out of another dataset. 

```

### Topological relations

(see [Geometric binary predicates on pairs of simple feature geometry sets](https://r-spatial.github.io/sf/reference/geos_binary_pred.html))

```{r}
# create a polygon
#
a_poly = st_polygon(list(rbind(c(-1, -1), c(1, -1), c(1, 1), c(-1, -1))))
a = st_sfc(a_poly)

# create a line
#
l_line = st_linestring(x = matrix(c(-1, -1, -0.5, 1), ncol = 2))
l = st_sfc(l_line)

# create points
#
p_matrix = matrix(c(0.5, 1, -1, 0, 0, 1, 0.5, 1), ncol = 2)
p_multi = st_multipoint(x = p_matrix)
p = st_cast(st_sfc(p_multi), "POINT")
```

```{r}
plot(a, col = "gray")
plot(l, add = TRUE)
plot(p, add = TRUE)
text(st_coordinates(p)[,1], st_coordinates(p)[,2], seq_along(p), pos = 4)
```

```{r}
st_intersects(p, a)
```

```{r}
st_disjoint(p, a)
```

```{r}
st_within(p, a)
```

```{r}
st_touches(p, a)
```

```{r}
st_is_within_distance(p, a, dist = 0.9)
```

### Spatial joining

```{r}
# random number generator seed
# (use constant for reproducibility)
#
set.seed(2018)

# world bounding box
#
(bb_world = st_bbox(world))

random_df = tibble(
  x = runif(n = 10, min = bb_world[1], max = bb_world[3]), 
  # bb_world[] ensures the points are in the world boundary box 
  y = runif(n = 10, min = bb_world[2], max = bb_world[4])
)

(random_points = random_df %>%
    st_as_sf(coords = c("x", "y")) %>%
    st_set_crs(4326))
```

```{r}
world_random = world[random_points,]
nrow(world_random)
```

```{r}
(random_joined = st_join(random_points, world["name_long"], left = FALSE))
```

```{r}
tmap_mode("view")
tm_shape(random_points) +
  tm_dots(col = "red", size = 0.1) +
  tm_shape(random_joined, size = 0.1) +
  tm_dots(col = "green", size = 0.1)
```

### Non-overlapping joins

```{r}
tmap_mode("view")

tm_shape(cycle_hire) +
  tm_dots(size = 0.05, alpha = 0.5, col = "blue") +
  tm_shape(cycle_hire_osm) +
  tm_dots(size = 0.05, alpha = 0.5, col = "red")

```

```{r}
any(st_touches(cycle_hire, cycle_hire_osm, sparse = FALSE))
```

```{r}
st_crs(cycle_hire)$Name
```

```{r}
st_crs(cycle_hire_osm)$Name
```

```{r}
st_crs(27700)$Name
```

```{r}
cycle_hire_P     = st_transform(cycle_hire,     27700)
cycle_hire_osm_P = st_transform(cycle_hire_osm, 27700)

sel = st_is_within_distance(cycle_hire_P, cycle_hire_osm_P, dist = 20)

summary(lengths(sel) > 0)
```

```{r}
z = st_join(cycle_hire_P, cycle_hire_osm_P,
            join = st_is_within_distance, dist = 20)

nrow(cycle_hire)
nrow(z)
```

```{r}
z = z %>% 
  group_by(id) %>% 
  summarize(capacity = mean(capacity))

nrow(z) == nrow(cycle_hire)
```

```{r}
plot(cycle_hire_osm["capacity"])
plot(z["capacity"])
```

### Spatial data aggregation

```{r}
nz_avheight = aggregate(x = nz_height, by = nz, FUN = mean)

nz_avheight2 = nz %>%
  st_join(nz_height) %>%
  group_by(Name) %>%
  summarize(elevation = mean(elevation, na.rm = TRUE))

tmap_mode("plot")

tm_shape(nz_avheight) +
  tm_fill(col = "elevation")

tm_shape(nz_avheight2) +
  tm_fill(col = "elevation")
```

### Distance relations

```{r}
nz_heighest = nz_height %>% top_n(n = 1, wt = elevation)
canterbury_centroid = st_centroid(canterbury)

st_distance(nz_heighest, canterbury_centroid)
```

```{r}
co = filter(nz, grepl("Canter|Otag", Name))
st_distance(nz_height[1:3, ], co)
```
